Baseline centering problem

### 2018-12-11 paper/tangotd.tex

Combined with all these problems, is the problem that a Bellman solution on the
support of $\rho$ may be very, very far from the global Bellman solution,
even on the support of $\rho$ (take random walk on a circle, with $\rho$
being everything but one point: then any linear function solves Bellman
on the support of $\rho$).

TODO seems to be much worse off-policy.

>>>>

Combined with all these problems, is the problem that an approximate
solution to the Bellman equation may be very, very far from the true
Bellman solution.  This is an absolutely major problem off-policy.

In a finite MDP, the solution to the Bellman equation is unique. However,
in infinite MDPs this is not the case at all. And even in finite MDPs,
the non-uniqueness problem becomes an ill-conditioning problem: some
functions are almost perfect solutions to the Bellman equation but can be
arbitrarily far from the true value function.

As an example, consider a one-dimensional problem with state $s\in \Z$.
Assume the policy $\pi$ can only jump to a nearest neighbor or stay in
place (with arbitrary probabilities). Then the Bellman equation on $V$
is, at each state $s$, a linear equation involving $V(s)$, $V(s+1)$ and
$V(s-1)$. It is easy to build many solutions by choosing arbitrary
values for $V(0)$ and $V(1)$, and then just solving the Bellman equation
in $V(s+1)$ for $s\geq 1$ and in $V(s-1)$ for $s\leq -1$.

For instance, imagine that the policy is a nearest-neighbor random walk
on $\Z$ (with probabilities $1/2$ to jump right or left), and that
$\gamma=1$. Then any linear function $f$ satisfies $f=Pf$, therefore,
linear functions lie in the kernel of the Bellman gap operator.
Consequently, if $V$ solves the Bellman equation, then $V+f$ is another
solution, for any linear function $f$.

What happens with linear functions is that the system ``believes''
there's a source of reward at infinity, and this belief cannot be
falsified via the Bellman equation alone.

(If the policy and environment are ergodic with invariant measure $\mu$,
and if rewards lie in $L^2(\mu)$, then the true value function is the
only solution of Bellman in $L^2(\mu)$, so penalizing the $L^2$-norm might
help a bit. However, some simple examples below are immune to this solution.)

Let us take another example. The state is $s\in \Z$. With nearest-neighbor transitions, the Bellman function of the optimal
policy satisfies (with $\gamma=1$)
\begin{equation}
\label{eq:Vopt}
V(s)=R(s)+\max_{s'\in \{s-1,s,s+1\}} V(s')
\end{equation}

For instance, if the reward is $0$ at $0$ and $-1$ everywhere else, the
optimal policy moves towards $0$, and the value function is
$V(s)=-\abs{s}$. However, it is easy to check that
$V(s)=\abs{s}+\1_{s=0}$ also solves the Bellman equation. This solution
believes in ``rewards at infinity''.

This ``almost'' happens in a finite state space. For instance, take a
state space $[-n;n]$ with reward $0$ at the boundaries $\pm n$, and
reward $-1$
everywhere else. The optimal policy goes to the nearest boundary point,
and the value function is $V(s)=\abs{s}-n$, shaped like a ``V''.

This solution is unique; however, there are near-solutions that violate
the Bellman equation only at one or a few points. For instance, $-V$ is a
solution everywhere except at $s=0$.

In an off-policy situation where we sample transitions $(s,s')$ and apply
TD at $(s,s')$, it would take a long time to realize that $-V$ is
not the correct solution, and convergence to the true solution would be
slow. This solution ``believes'' there is a good reward at $s=0$, and
this belief is consistent as long as we do not actually reach $s=0$.

This phenomenon is fully general and not specific to 1D problems; in any
MDP one can build near-Bellman solutions
that ``believe'' there is a reward at a state $s_0$ and solve Bellman
exactly everywhere except at $s_0$. Off-policy these violations would not be
detected. On-policy, the policy would move towards $s_0$ and detect the
problem. I believe this is a serious limitation of off-policy approaches.

Are these pseudo-solutions frequent, or does a good initialization move
away from such solutions?
The continuous-state limit is interesting for this.

We can view $[-n;n]$ as a
discretization of a problem with state $s\in [-1;1]$, where the state is
discretized with step $\eps=1/n$. Then the continuous-state limit of the
Bellman equation \eqref{eq:Vopt} for the optimal policy (with $\gamma=1$)
is just
\begin{equation}
\abs{\partial V(s)/\partial s}=-R(s)
\end{equation}
(as one can see by expanding $V(s+\eps)$ and $V(s-\eps)$ in the equation
$V(s)=R(s)\eps +\max \{V(s+\eps),V(s-\eps)\}$; for simplicity I have
discarded the possibility to stay in place). [One gets a constraint that
rewards are negative because in such a situation, the optimal policy
moves to the state with highest reward, so with $\gamma=1$ the value
function is well-defined only if one shifts the rewards by the maximum
reward so that the maximum reward becomes $0$.]

This equation is wildly underdetermined. For instance, if $V$ is a
solution then $-V$ is also a solution...
